---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

Set WD
```{r}
setwd("~/R/Coursera/05 Rep/Assignment1")
```

Read CSV
```{r}
data <- read.csv(unz("activity.zip", "activity.csv"), colClasses=c("integer", "Date", "integer"))
data$interval <- 60*floor((data$interval+1)/100) + (data$interval %% 100)
str(data)
summary(data)
```

## What is mean total number of steps taken per day?
Average number of steps per day
```{r}
step_daily_sum <- tapply(data$steps, data$date, sum, na.rm=T)
step_mean <- mean(step_daily_sum)
step_mean
```
Median number of steps per day
```{r}
step_median <- median(step_daily_sum)
step_median
```
Histogramm Median vs. Mean

```{r}
hist(step_daily_sum, breaks=11, 
     xlab="number of steps", 
     main="Histogram of steps per day")
abline(v=step_mean, col="cyan", lwd=3)
abline(v=step_median, col="magenta", lwd=3)
legend(x="topright", legend=c("Mean","Median"), col=c("cyan","magenta"), bty="n", lwd=3)
```


## What is the average daily activity pattern?
```{r}
step_avg <- tapply(data$steps, data$interval, mean, na.rm=T)

hours <- as.numeric(names(step_avg))/60
plot(hours, step_avg, type="l", axes=F,
     xlab="time of day (h)", ylab="average number of steps in 5-min interval",
     main="Daily Activity Pattern")
axis(2)
axis(1, at=0:6*4, labels=paste(0:6*4,":00", sep=""))
```

interval with max number of steps
```{r}
maxact <- which(step_avg==max(step_avg))
maxint <- data$interval[maxact]
sprintf("%02d:%02d", floor(maxint/60), maxint %% 60)
```

## Imputing missing values
Number of NAs
```{r}
sum(is.na(data))
```
Impute
```{r}
impute <- transform(data, steps=ifelse(is.na(steps), step_avg, steps))

summary(impute)
```

```{r}
total_impsteps <- tapply(impute$steps, impute$date, sum, na.rm=T)

impstep_mean <- mean(total_impsteps)
impstep_mean

impstep_median <- median(total_impsteps)
impstep_median

```

Histogram
```{r}
hist(total_impsteps, breaks=11, 
     xlab="number of steps", 
     main="Histogram of steps per day")
abline(v=impstep_mean, col="cyan", lwd=3,lty=1)
abline(v=impstep_median, col="magenta", lwd=3,lty=3)
legend(x="topright", legend=c("Mean","Median"), col=c("cyan","magenta"), bty="n", lwd=3)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
Sys.setlocale("LC_TIME", "English")

week <- factor(weekdays(impute$date) %in% c("Saturday","Sunday"), 
               labels=c("Weekday","Weekend"), ordered=FALSE)

step_imp <- aggregate(impute$steps, by=list(interval=impute$interval, weekday=week), mean)

library(ggplot2)
g <- ggplot(step_imp, aes(interval/60, x))
g + geom_line() + facet_grid(weekday ~ .) +
    scale_x_continuous(breaks=0:6*4, labels=paste(0:6*4,":00", sep="")) +
    theme_bw() +
    labs(y="average number of steps per 5-minute interval") +
    labs(x="daytime") +
    labs(title="Daily activity pattern - Weekday vs Weekend")
               
```

